Cover Letter: Cudele

### 2018-01-07 paper/cover_letter.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

We thank the reviewers for their hard work and genuinely helpful suggestions.
In addition to this cover letter, we have posted a PDF online explicitly
showing additions in \textcolor{red}{\textbf{red}} and deletions in
\textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
feedback; Reviewer 1 and 3's comments are addressed in issues II and III,
Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
comments are addressed in issues I and II.

@@

We re-wrote Section~{\S}V-B to highlight examples from HPC and cloud workloads
that would benefit from using Cudele. For HPC we focus on user home directories
used for experiments and for the cloud we focus on the Hadoop and Spark
runtimes (as suggested by Reviewer 2).  We also added \textbf{In Cudele}
headers in Section~{\S}V-B to show which subtree semantics accommodate each
workload and to demonstrate how Cudele might be able to help in these
scenarios.  By showing these setups, we hope to demonstrate that different
applications can use a file system in parallel with different
consistency/durability semantics.  Although we do not actually run the
workloads on our prototype, hopefully our changes show how the use cases, and
the synthetic benchmarks we used to represent them, are indeed relevant to
parallel and distributed computing.

To align the paper to the themes of the conference, we changed Section~{\S}V-A
and Figures 6a, 6b, and 6c to emphasize that decoupled namespaces facilitate
client-driven parallelism, where clients detach subtrees form the global namespaces
and do operations in-parallel to their local disk. We also hope that changes to
the use cases above motivates the need for robust distributed storage systems
that can handle today's applications.  So while the paper is storage-centric,
we try to show that the workloads it supports are highly distributed and need a
flexible solution like Cudele.

@@

We apologize for the omission of experimental setup and source code details.
We made artifacts available online for all experiments in the paper. For the
experimental setup, we describe the cluster setup (hardware, software, etc.) in
Section~{\S}V.  We have also made the infrastructure code available and added
links after each figure to show exactly how experiments are run.  This
infrastructure code contains scripts to deploy the system, run experiments, and
gather results.  This process follows the Popper
Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
aims to make research reproducible.  We also made the source code available
online and a link is provided in a reference in Section~{\S}V. Finally, we add
details about which tools and classes we modified in Section~{\S}IV to address
Reviewer 1's questions about the framework implementation.

%\section*{Cosmetic Fixes}
%%Terminology
%%- users vs. clients R2
%\begin{itemize}
%  \item Reviewer 1: fixed user vs. client terminology in Section~{\S}III
%  \item Reviewer 1: clarified that the cost of dynamically changing semantics is future work Sections~{\S}III and ~{\S}V.E
%  \item Reviewer 1: added source code pointer in Section~{\S}IV
%  \item Reviewer 2: remove major takeways in Section~{\S}V
%  \item Reviewer 2: remove major cross-references in Section~{\S}V
%  \item Reviewer 3: quantified speedups with figure annotations in Section~{\S}V
%  \item Reviewers 1 and 3: add servers, network, and storage setups in Section~{\S}V
%  \item Reviewer 4: add new section ({\S}V.F) describing how Cudele would work with a system like Spark
%\end{itemize}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,paper}
\end{document}

>>>>

We thank the reviewers for their hard work and genuinely helpful suggestions.
In addition to this cover letter, we have posted a
\href{https://github.com/michaelsevilla/cudele-popper/blob/master/paper/diff.pdf}{PDF
online} explicitly showing additions in \textcolor{blue}{\textbf{blue}} and
deletions in \textcolor{red}{\textbf{red}}. Three issues were common across
reviewer feedback; Reviewer 1 and 3's comments are addressed in issues II and
III, Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
comments are addressed in issues I and II.

@@

We re-wrote Section~{\S}V-B to highlight examples from HPC and cloud workloads
that would benefit from using Cudele. For HPC we focus on user home directories
used for experiments and for the cloud we focus on the Hadoop and Spark
runtimes (as suggested by Reviewer 2).  We also added \textbf{Cudele Setup}
headers in Section~{\S}V-B to show which subtree semantics accommodate each
workload and to demonstrate how Cudele might be able to help in these
scenarios.  By showing these setups, we hope to demonstrate that different
applications can use a file system in parallel with different
consistency/durability semantics.  Although we do not actually run the
workloads on our prototype, hopefully our changes show how the use cases, and
the synthetic benchmarks we used to represent them, are indeed relevant to
parallel and distributed computing.

To align the paper to the themes of the conference, we changed Section~{\S}V-A
and Figures 6a, 6b, and 6c to emphasize that decoupled namespaces facilitate
client-driven parallelism, where clients detach subtrees form the global namespaces
and do operations in-parallel to their local disk. We also hope that changes to
the use cases above motivate the need for robust distributed storage systems
that can handle today's applications.  So while the paper is storage-centric,
we try to show that the workloads it supports are highly distributed and need a
flexible solution like Cudele.

@@

We apologize for the omission of experimental setup and source code details.
We made artifacts available online for all experiments in the paper. For the
experimental setup, we describe the cluster setup (hardware, software, etc.) in
Section~{\S}V.  We have also made the infrastructure code available and added
links after each figure to show exactly how experiments are run.  This
infrastructure code contains scripts to deploy the system, run experiments, and
gather results.  This process follows the \href{http://falsifiable.us/}{Popper
Convention}, which aims to make research reproducible.  We also made the source
code available online and a link is provided in Section~{\S}V. Finally, we add
details about which tools and classes we modified in Section~{\S}IV to address
Reviewer 1's questions about the framework implementation.

\section*{Additional Cosmetic Fixes}
%%Terminology
%%- users vs. clients R2

For Reviewer 1, we fixed user vs. client terminology in Section~{\S}III,
clarified that the cost of dynamically changing semantics is future work in
Sections~{\S}I and~{\S}III, and added source code pointers in Section~{\S}V.
For Reviewer 2, we removed the major takeways headings and cross-references in
Section~{\S}V. For Reviewer 3, we quantified speedups with figure annotations
in Section~{\S}V. For Reviewers 1 and 3, we added servers, network, and storage
setups in Section~{\S}V. For Reviewer 4 we added descriptions to each use case
in Section~{\S}V describing how Cudele would work with Spark.

\end{document}

### 2018-01-07 paper/sections/0_abstract.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

HPC and data center scale application developers are abandoning POSIX IO
because file system metadata synchronization and serialization overheads of
providing strong consistency and durability are too costly -- and often
unnecessary -- for their applications.  Unfortunately, designing file systems
with weaker consistency or durability excludes applications that rely on
stronger guarantees, forcing developers to re-write their applications or
deploy them on a different system.  We present a framework and API that lets
\oldcomment{clients}\newcomment{administrators} specify their
consistency/durability requirements and dynamically assign them to subtrees in
the same namespace, allowing administrators to optimize subtrees over time and
space for different workloads. \oldcomment{By custom fitting a subtree to a
create-heavy application,} We show similar speedups to related work but more
importantly, \oldcomment{our prototype can custom fit subtrees in the same
namespace to applications common in large data centers, }\newcomment{ we show
performance improvements when we custom fit subtree semantics to applications}
such as checkpoint-restart (91.7\(\times\) speedup), user home directories
(0.03 standard deviation from optimal), and
\oldcomment{clients}\newcomment{users} checking for partial results (2\%
overhead). \oldcomment{ and can scale to 2\(\times\) as many clients when
compared to our baseline system.}

>>>>

HPC and data center scale application developers are abandoning POSIX IO
because file system metadata synchronization and serialization overheads of
providing strong consistency and durability are too costly -- and often
unnecessary -- for their applications.  Unfortunately, designing file systems
with weaker consistency or durability excludes applications that rely on
stronger guarantees, forcing developers to re-write their applications or
deploy them on a different system.  We present a framework and API that lets
\oldcomment{clients}\newcomment{administrators} specify their
consistency/durability requirements and dynamically assign them to subtrees in
the same namespace, allowing administrators to optimize subtrees over time and
space for different workloads. \oldcomment{By custom fitting a subtree to a
create-heavy application,} We show similar speedups to related work but more
importantly, \oldcomment{our prototype can custom fit subtrees in the same
namespace to applications common in large data centers, }\newcomment{ we show
performance improvements when we custom fit subtree semantics to applications
such as checkpoint-restart (91.7\(\times\) speedup), user home directories
(0.03 standard deviation from optimal), and users checking for partial results
(2\% \oldcomment{clients}\newcomment{users} overhead).} \oldcomment{ and can
scale to 2\(\times\) as many clients when compared to our baseline system.}

### 2018-01-07 paper/sections/3_posix_overheads.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb] \centering
\includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
\caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-compile/visualize/viz.ipynb}{source}]
For the CephFS metadata server, create-heavy workloads ({\it e.g.},
\texttt{untar}) incur the highest disk, network, and CPU utilization because of
consistency/durabiliy demands.}\label{fig:overhead-creates}
\end{figure}

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.32\linewidth}
      \centering
      \includegraphics[width=1\linewidth]{graphs/slowdown-journal.png}
      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-durability/visualize/viz.ipynb}{source}]
      effect of journal} \label{fig:overhead-a}
  \end{subfigure}
  \begin{subfigure}[b]{.32\linewidth}
      \centering
      \includegraphics[width=0.98\linewidth]{graphs/slowdown-interfere-scale.png}
      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-creates/visualize/viz.ipynb}{source}]
      interference hurts variability}
      \label{fig:overhead-b}
  \end{subfigure}
  \begin{subfigure}[b]{.32\linewidth}
      \centering
      \includegraphics[width=1.15\linewidth]{graphs/behavior-interfere.png}
      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-interfere/visualize/viz.ipynb}{source}]
      interference increases RPCs}
      \label{fig:overhead-c}
  \end{subfigure}
  \caption{\newcomment{The durability and strong consistency slowdown of the
existing CephFS implementation increases as the number of clients scales.
Results in (a) and (b)  are normalized to 1 client that creates 100K files in
isolation.  (a) shows the effect of journaling metadata updates; ``segment(s)"
is the number of journal segments dispatched to disk at once. (b) shows the
slowdown when another client interferes by creating files in all directories
and (c) highlights the cause: }\oldcomment{The overhead of durability and
strong consistency in CephFS. (a) shows the effect of different journal segment
sizes, which are streamed into the object store for fault tolerance. (b) and
(c) show that} when another client ``interferes", capabilities are revoked and
metadata servers do more work.} \label{fig:overhead}
\end{figure*}

>>>>

\begin{figure}[tb] \centering
\includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
\caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-compile/visualize/viz.ipynb}{source}]
For the CephFS metadata server, create-heavy workloads ({\it e.g.},
\texttt{untar}) incur the highest disk, network, and CPU utilization because of
consistency/durabiliy demands.}\label{fig:overhead-creates}
\end{figure}

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.32\linewidth}
      \centering
      \includegraphics[width=1\linewidth]{graphs/slowdown-journal.png}
      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-durability/visualize/viz.ipynb}{source}]
      effect of journal} \label{fig:overhead-a}
  \end{subfigure}
  \begin{subfigure}[b]{.32\linewidth}
      \centering
      \includegraphics[width=0.98\linewidth]{graphs/slowdown-interfere-scale.png}
      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-creates/visualize/viz.ipynb}{source}]
      interference hurts variability}
      \label{fig:overhead-b}
  \end{subfigure}
  \begin{subfigure}[b]{.32\linewidth}
      \centering
      \includegraphics[width=1.15\linewidth]{graphs/behavior-interfere.png}
      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-interfere/visualize/viz.ipynb}{source}]
      interference increases RPCs}
      \label{fig:overhead-c}
  \end{subfigure}
  \caption{\newcomment{The durability and strong consistency slowdown of the
existing CephFS implementation increases as the number of clients scales.
Results in (a) and (b)  are normalized to 1 client that creates 100K files in
isolation.  (a) shows the effect of journaling metadata updates; ``segment(s)"
is the number of journal segments dispatched to disk at once. (b) shows the
slowdown when another client interferes by creating files in all directories
and (c) highlights the cause: }\oldcomment{The overhead of durability and
strong consistency in CephFS. (a) shows the effect of different journal segment
sizes, which are streamed into the object store for fault tolerance. (b) and
(c) show that} when another client ``interferes", capabilities are revoked and
metadata servers do more work.} \label{fig:overhead}
\end{figure*}

### 2018-01-07 paper/sections/6_results.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\newcomment{\indent We graph standard deviations for three runs (sometimes
error bars are too small to see) and normalize results to make our results more
generally applicable to different hardware. We use a CloudLab cluster of 34
nodes connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and
400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
release, version 10.2.1, which was released in May 2016. All daemons run as a
single process (and the Ceph components~\cite{docs:cephrados} are: 1 monitor, 3
object storage daemons, 1 metadata server, and 20 clients.} We scope the
evaluation to one metadata server and scale the number of parallel clients
\newcomment{each doing 100K operations because 100K is the maximum recommended
size of a directory in CephFS. We scale to 20 clients because, as shown in
Section~\S\ref{sec:posix-overheads}, 20 clients is enough to saturate the
resources of a single metadata server.} This type of analysis shows the
capacity and performance of a metadata server with superior metadata protocols,
which should be used to inform \oldcomment{load balancing}\newcomment{metadata
distribution} across a \oldcomment{metadata} cluster. Load balancing across a
cluster of metadata servers with partitioning and replication can be explored
with something like Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our
results reproducible, this paper adheres to The Popper
Convention~\cite{jimenez_popper_2016}, so experiments can be examined in more
detail, or even re-run, by visiting the \texttt{[source]} link next to each
figure.  The source code is available on the Cudele branch~\cite{cudele-src} of
our Ceph fork.}

@@

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
\newcomment{Overhead of processing 100K create events for each mechanism in
\oldcomment{Table~\ref{table:mechanisms}}\newcomment{Figure~\ref{fig:decouple}},
normalized to the runtime of writing events to client memory. The far right
graph shows the overhead of building semantics of real world systems.}
\oldcomment{Performance of each mechanism (left) and building the
consistency/durability semantics of real-world systems (right) for 100K files
creates from a single client. Results are normalized to the runtime of writing
events to the client's in-memory journal.}\label{fig:composable-mechanisms}}
\end{figure}

@@

\begin{figure*}[tb]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
      \newcomment{parallel creates on clients}
      }\label{fig:mergescale}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
      \newcomment{block interference}
      }\label{fig:block-allow}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
      \newcomment{syncing to global namespace}
      }\label{fig:slowdown-sync}
  \end{subfigure}
\caption{\oldcomment{ The performance and features of Cudele}\newcomment{Cudele
performance}. (a) \newcomment{shows the speedup of decoupled namespaces over
RPCs; \texttt{create} is the throughput of clients creating files in-parallel
and writing updates locally; \texttt{create+merge} includes the time to merge
updates at the metadata server. }\oldcomment{shows the cost of merging client
journals at the metadata server. Shipping and merging journals of
updates}\newcomment{Decoupled namespaces} scale better than RPCs because there
are less messages and consistency/durability code paths are bypassed. (b) shows
how the allow/block API isolates directories from interfering clients. (c) is
the slowdown of a single client syncing updates to the global namespace. The
inflection point is the trade-off of frequent updates vs. larger journal
files.}
\label{fig:use-cases}
\end{figure*}

>>>>

\newcomment{\indent We graph standard deviations for three runs (sometimes
error bars are too small to see) and normalize results to make our results more
generally applicable to different hardware. We use a CloudLab cluster of 34
nodes connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and
400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
release, version 10.2.1, which was released in May 2016. We use 1 monitor
daemon, 3 object storage daemons, 1 metadata server daemon, and up to 20
clients.}  We scope the evaluation to one metadata server and scale the number
of parallel clients \newcomment{each doing 100K operations because 100K is the
maximum recommended size of a directory in CephFS. We scale to 20 clients
because, as shown in Section~\S\ref{sec:posix-overheads}, 20 clients is enough
to saturate the resources of a single metadata server.} This type of analysis
shows the capacity and performance of a metadata server with superior metadata
protocols, which should be used to inform \oldcomment{load
balancing}\newcomment{metadata distribution} across a \oldcomment{metadata}
cluster. Load balancing across a cluster of metadata servers with partitioning
and replication can be explored with something like
Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our results
reproducible, this paper adheres to The Popper
Convention~\cite{jimenez_popper_2016}, so experiments can be examined in more
detail, or even re-run, by visiting the \texttt{[source]} link next to each
figure.  The source code is available on the Cudele branch~\cite{docs:cudele}
of our Ceph fork.}

@@

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
\newcomment{Overhead of processing 100K create events for each mechanism in
\oldcomment{Table~\ref{table:mechanisms}}\newcomment{Figure~\ref{fig:decouple}},
normalized to the runtime of writing events to client memory. The far right
graph shows the overhead of building semantics of real world systems.}
\oldcomment{Performance of each mechanism (left) and building the
consistency/durability semantics of real-world systems (right) for 100K files
creates from a single client. Results are normalized to the runtime of writing
events to the client's in-memory journal.}\label{fig:composable-mechanisms}}
\end{figure}

@@

\begin{figure*}[tb]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
      \newcomment{parallel creates on clients}
      }\label{fig:mergescale}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
      \newcomment{block interference}
      }\label{fig:block-allow}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
      \newcomment{syncing to global namespace}
      }\label{fig:slowdown-sync}
  \end{subfigure}
\caption{\oldcomment{ The performance and features of Cudele}\newcomment{Cudele
performance}. (a) \newcomment{shows the speedup of decoupled namespaces over
RPCs; \texttt{create} is the throughput of clients creating files in-parallel
and writing updates locally; \texttt{create+merge} includes the time to merge
updates at the metadata server. }\oldcomment{shows the cost of merging client
journals at the metadata server. Shipping and merging journals of
updates}\newcomment{Decoupled namespaces} scale better than RPCs because there
are less messages and consistency/durability code paths are bypassed. (b) shows
how the allow/block API isolates directories from interfering clients. (c) is
the slowdown of a single client syncing updates to the global namespace. The
inflection point is the trade-off of frequent updates vs. larger journal
files.}
\label{fig:use-cases}
\end{figure*}

### 2018-01-07 paper/sections/7_conclusion.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Relaxing consistency/durability semantics in file systems is a double-edged
sword. While it performs and scales better, it alienates applications that rely
on strong consistency and durability.  Cudele lets
\oldcomment{users}\newcomment{administrators} assign consistency/durability
guarantees to subtrees in the global namespace, which lets administrators
custom fit semantics to application. We show how applications can co-exist and
perform well in a global namespace and our prototype enables studies that
adjust these semantics over {\it time and space}, where subtrees can change
without ever moving the data they reference.

%This approach is superior to mounting different storage systems in a global
%namespace because there is no need to provision dedicated storage clusters to
%applications or move data between these systems.  
Cudele prompts many avenues for future work.  First is dynamically changing
semantics of a subtree from stronger to weaker guarantees (or vice versa).  For
example, the results of a Hadoop job do not need to be migrated into CephFS for
other processing; instead the \oldcomment{user}\newcomment{administrator} can
change the semantics of the HDFS subtree into a CephFS subtree, which may cause
metadata/data movement to make ensure strong consistency. Second is embeddable
policies, where child subtrees have specialized features but still maintain
guarantees  of their parent subtrees.  For example, a RAMDisk subtree is POSIX
IO-compliant but relaxes durability constraints, so it can reside under a POSIX
IO subtree that is strongly consistent.

>>>>

Relaxing consistency/durability semantics in file systems is a double-edged
sword. While it performs and scales better, it alienates applications that rely
on strong consistency and durability.  Cudele lets
\oldcomment{users}\newcomment{administrators} assign consistency/durability
guarantees to subtrees in the global namespace, resulting in 
custom fit semantics for application. We show how applications can co-exist and
perform well in a global namespace and our prototype enables studies that
adjust these semantics over {\it time and space}, where subtrees can change
without ever moving the data they reference.

%This approach is superior to mounting different storage systems in a global
%namespace because there is no need to provision dedicated storage clusters to
%applications or move data between these systems.  
Cudele prompts many avenues for future work.  First is to co-locate HPC
workflows with real highly parallel runtimes from the cloud in the same
namespace. This setup shows how Cudele reasonably incorporates both programming
models (client-driven parallelism and user-defined workflows) at the same time
and should show large performance gains.  Second is dynamically changing
semantics of a subtree from stronger to weaker guarantees (or vice versa).  For
example, the results of a Hadoop job do not need to be migrated into CephFS for
other processing; instead the \oldcomment{user}\newcomment{administrator} can
change the semantics of the HDFS subtree into a CephFS subtree, which may cause
metadata/data movement to ensure strong consistency. Third is embeddable
policies, where child subtrees have specialized features but still maintain
guarantees of their parent subtrees.  For example, a RAMDisk subtree is POSIX
IO-compliant but relaxes durability constraints, so it can reside under a POSIX
IO subtree that is strongly consistent.
